A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-21 doc/paper.tex
Fixes #46

		\item [1] \textbf{Provide an interactive environment for unguided exploration.}

		%The combination space for all Excel functions is massive, let
		%alone the space for observed formulas. Working from data with actual referents
		%in practice, the tool must aid the user in navigating this space.

		\item [!1] \textbf{Do not make judgments about good and bad design.}

		%Though
		%the tool tries to foster tries understanding by linking pattern to example, it
		%won't provide a precise description of what a function accomplishes. The tool's
		%user must infer this.

		\item [2] \textbf{Emphasize the quantitative patterns in formula construction.}

		%The tool, accommodating datasets of a few spreadsheets to millions, should
		%address the questions of how often the end users employed a certain function
		%and where they used it. In this way, it must show precise metrics, such as
		%frequency of  use and depth of function nesting, of the dataset it conveys.

		\item [!2] \textbf{Do not overwhelm with numbers.}

		%This is essentially an
		%exploratory tool, not a generative one. It should not produce any information
		%other than new views of the original data.

		\item [3] \textbf{Promote a qualitative understanding of the patterns.}

		%Lest
		%these patterns remain abstract, the tool should supplement its observations
		%with concrete instances of relevant formulas from the corpus. Ideally, it
		%should even direct users to the exact cell in the spreadsheets where the
		%formula was used, contextualizing the functions.

		\item [!3] \textbf{Do not attempt to explain formulae.}

		%Though there is room
		%to explore the place of a single function in the group, the tool must design
		%the core visualization around the entire body of data, not the other way
		%around.

@@

	\subsection{Walkthrough of Tool} Before discussing the minutiae of how the tool
	was developed, it will help first to give a basic example of how the tool can
	be used and how to interpret its symbols. One such example might be to find the
	answer to this question: \textit{ What kinds of functions do people use to
		define the condition in an IF function?} The final visualization is available
	in figure \ref{fig:fullpic}; however, to understand the interaction, we will
	start rather from the beginning. \par

@@

	\centerline{\includegraphics[width=.43\textwidth]{IFexpand}}

@@

	\subsection{Design Decisions} Early in the process, we decided that the tree
	form would be a suitable fit to represent the parent-child or caller-callee
	relationships inherent in the data, given the composition of formulas as
	functions and their arguments (which could be yet more functions). By adapting
	this structure to accommodate a broad range of possibilities for nested
	functions, the branching factor depends on both the number of arguments in a
	function and the number of possible functions observed as an argument in a
	function. \par

>>>>

		\item [1] \textbf{Provide an interactive environment for free exploration.}
		There are a lot of data in some of these collections, and the user should be
		able to pick for themselves which to explore while ignoring the rest. Furthermore,
		this exploration should not be designed around answering one specific question
		but rather be open-ended.

		\item [!1] \textbf{Do not make judgments about good and bad design.} Though
		users may certainly approach the tool with the intention of finding particular
		formulae, \toolname will not prioritize any on metrics of good or bad
		practice. Rather, it will present them agnostically, referring only to
		frequency.

		\item [2] \textbf{Emphasize the quantitative patterns in formula
			construction.} \toolname should address the questions of how often the
		spreadsheet programmers used a certain function and where they used it. In
		this way, it must show precise metrics, such as frequency of use and depth of
		function nesting, of the dataset it conveys.

		\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		especially, we anticipated some particularly dense trees. Though
		everything should be accessible in accordance with Goal 1, we should design in
		a way that reduces noise.

		\item [3] \textbf{Promote a qualitative understanding of the patterns.}
		\toolname should supplement its observations with concrete instances of
		relevant formulas from the corpus. Ideally, it should even direct users to the
		exact cell in the spreadsheets where the formula was used, contextualizing the
		functions.

		\item [!3] \textbf{Do not attempt to explain formulae.} Though \toolname tries
		to foster understanding by linking pattern to example, it should not provide
		a precise description of what a complex formula accomplishes. The tool's user
		must infer this on their own.

@@

	Though we kept these guidelines in mind through the development of \toolname, we
	nevertheless reached a number of points in which these principles conflicted. For
	a discussion of these decisions, see section \ref{ssec:decisions}.

	\subsection{Walkthrough of the Tool} Before discussing the minutiae of how the tool
	was developed, it will help first to give a basic example of how the tool can
	be used and how to interpret its symbols. One such example might be to find the
	answer to this question: \textit{ What kinds of functions do people use to
		define the condition in an IF function?} The final visualization is available
	in figure \ref{fig:fullpic}; however, to understand the interaction, we will
	start rather from the beginning. \par

@@

	\begin{figure}
		\centering
		\includegraphics[width=.43\textwidth]{IFexpand}
		\label{fig:expandif}
		\caption{First argument of the IF tree}
	\end{figure}

@@

	\subsection{Design Decisions} \label{ssec:decisions}
	Early in the process, we decided that the tree
	form would be a suitable fit to represent the parent-child or caller-callee
	relationships inherent in the data, given the composition of formulas as
	functions and their arguments (which could be yet more functions). By adapting
	this structure to accommodate a broad range of possibilities for nested
	functions, the branching factor depends on both the number of arguments in a
	function and the number of possible functions observed as an argument in a
	function. \par
